---
title: "Test with an Artificial Sinusoid Signal"
output: html_notebook
---

Create an artificial signal with three sinusoid components and gaussian noise, and sample it at regular intervals:

```{r}
# Required library
library("multitaper")

# Artificial signal with frequencies 3, 4, 4.5 sampled at regular times
t <- seq(1, 10, length.out = 1000)
sig <- sin(2*pi*3*t) + sin(2*pi*4*t) + sin(2*pi*4.5*t)
N <- length(t)
T <- diff(range(t))

# Add noise to signal
noisy.sig <- sig + rnorm(length(sig), mean=0, sd=0.5)
```

Plot the signal:

```{r}
# Plot signal
plot(t, sig, type='l', xlab = 'time', ylab = 'Signal', main = 'Artificial signal')
plot(t, noisy.sig, type='l', xlab = 'time', ylab = 'Signal', main = 'Artificial signal w noise')
```

Use the ``multitaper`` package to estimate the multitaper spectrum given even-sampling.

```{r}
# Multitaper Spectral Estimate of signal - even sampling
sigMT <- spec.mtm(sig, nw=4.0, k=7, deltat = T/N, Ftest = TRUE)
sigMTnoise <- spec.mtm(noisy.sig, nw=4.0, k=7, deltat = T/N, Ftest = TRUE)

# Plot Ftest statistic
plot(sigMT$freq, sigMT$mtm$Ftest, type = 'l', main = 'F test for signal',
     xlab = 'Frequency', ylab = 'F statistic')
plot(sigMTnoise$freq, sigMTnoise$mtm$Ftest, type = 'l', main = 'F test for noisy signal',
     xlab = 'Frequency', ylab = 'F statistic', xlim = c(0, 20))
```

Use the ``mtls`` package to estimate the multitaper LS periodogram. Since we use evenly-sampled data, the ``mtls`` should perform the same as ``multitaper``.

```{r}
# Multitaper LS Periodogram of signal
sigMTLS <- LSspecMT(t, sig, w = 4.0/N, k = 7, Ftest=TRUE)
sigMTLSnoise <- LSspecMT(t, noisy.sig, w = 4.0/N, k = 7, Ftest=TRUE)

# Plot spectrum and Ftest
plot(sigMTLS, type = 'l', main = 'MTLS for signal')
sigMTLSFtest <- attr(sigMTLS, 'Ftest')
plot(sigMTLS$freq, sigMTLSFtest$F_, type = 'l', main = 'MTLS F test for signal',
     xlab = 'Frequency', ylab = 'F statistic')

# Plot spectrum and Ftest zoomed
plot(sigMTLS, type = 'l', main = 'MTLS (zoomed)', xlim = c(0, 20))
plot(sigMTLS$freq, sigMTLSFtest$F_, type = 'l', main = 'MTLS F test (zoomed)',
     xlab = 'Frequency', ylab = 'F statistic', xlim = c(0, 20))

sigMTLSnoiseFtest <- attr(sigMTLSnoise, 'Ftest')
plot(sigMTLSnoise, type = 'l', main = 'MTLS noisy signal')
# Plot Ftest
plot(sigMTLSnoise$freq, sigMTLSnoiseFtest$F_, type = 'l', main = 'MTLS F test noisy signal',
     xlab = 'Frequency', ylab = 'F statistic')
```
```{r}
# Non-uniform discrete fourier transform
sigNUMT <- NUspecMT(t, sig, w = 4.0/N, k = 7, Ftest = TRUE)

# Plot NUDFT periodogram and Ftest
plot(sigNUMT, type = 'l', main = 'NUMT for signal')
# Plot Ftest
sigNUMTFtest <- attr(sigNUMT, 'Ftest')
plot(sigNUMT$freq, sigNUMTFtest$F_, type = 'l', main = 'NUMT F test for signal',
     xlab = 'Frequency', ylab = 'F statistic')

# Plot NUDFT periodogram
plot(sigNUMT, type = 'l', main = 'NUMT (zoomed)', xlim = c(0, 20))
# Plot Ftest
plot(sigNUMT$freq, sigNUMTFtest$F_, type = 'l', main = 'NUMT F test (zoomed)',
     xlab = 'Frequency', ylab = 'F statistic', xlim = c(0, 20))
```
Test using ``willamete`` evenly-sampled data:

```{r}
# Willamete dataset
data(willamette)
t <- seq(1, 100, length.out = length(willamette))
N <- length(t)
T <- diff(range(t))

# Plot signal
plot(t, willamette, type='l', main = 'Willamete data', ylab = 'Signal')
```

```{r}
# Multitaper spectral estimate
willMT <- spec.mtm(willamette, nw=4.0, k=7, deltat=T/N, Ftest = TRUE)
plot(willMT$freq, willMT$mtm$Ftest, type = 'l', main = 'MTM F test',
     xlab = 'Frequency', ylab = 'F statistic')

# MTLS
willMTLS <- LSspecMT(t, willamette, w= 4.0/N, k = 7)
plot(willMTLS, type = 'l', main = 'MTLS Periodogram')
willMTLSFtest <- attr(willMTLS, 'Ftest')
plot(willMTLS$freq, willMTLSFtest$F_, type = 'l', main = 'MTLS F test',
     xlab = 'Frequency', ylab = 'F statistic')

# NUDFT
willNMT <- NUspecMT(t, willamette, w = 4.0/N, k = 7, Ftest = TRUE)
plot(willNMT, type = 'l', main = 'NUMT Periodogram')
willNMTFtest <- attr(willNMT, 'Ftest')
plot(willNMT$freq, willNMTFtest$F_, type = 'l', main = 'NUMT F test',
     xlab = 'Frequency', ylab = 'F statistic')
```

